An Open Repository of Real-Time COVID-19 Indicators

### 2021-10-04 indicators/paper/paper.tex
Edit text around Fig 6

Since July 2020, we have been regularly submitting short-term forecasts of
COVID-19 case and death incidence, at the state and county levels, to the
COVID-19 Forecast Hub \cite{ForecastHub}, with ``CMU-TimeSeries'' as the
team-model name.  The process of building, training, and deploying our
forecasting models leverages much of the infrastructure described in this paper
(such as the COVIDcast API's \textit{as of} feature), and some of our
forecasting systems rely on auxiliary indicators (such as survey-based and
claims-based COVID-like illness signals, which are described below).

@@

Besides tracking contemporaneous COVID activity, these and other indicators can
be used to improve forecasts of future COVID case trends, as investigated in a
companion paper \cite{McDonald:2021}.

@@

Figure~\ref{fig:geo_wise_correlation} shows the geo-wise correlations achieved
by the five signals and COVID case rates (from JHU CSSE, smoothed using a 7-day
trailing average), from April 15, 2020 to April 15, 2021. This calculation is
performed over all counties with at least 500 cumulative cases by the end of
this period, and at which all indicators are available (956 counties in total).
The large positive correlations suggest that these signals could be useful in
hotspot detection (identifying counties that have relatively high COVID
activity, at a given time). Somewhat surprisingly, the survey-based
CLI-in-community signal shows the strongest correlations for much of the time
period. This clearly demonstrates the value of a large-scale survey such as CTIS
for tracking symptoms and case trends, especially when other data is
unavailable. The correlations vary over time in complex ways; while some of this
variation is doubtless due to complex changes in public behavior, reporting and
testing practices, and so on, some is also related to changes in overall
COVID-19 case rates. For example, comparing against
Figure~\ref{fig:time_trends_national}, we observe that correlations drop in
February 2021 for several of the signals, matching when COVID cases also sharply
decline. This decline likely caused the heterogeneity in case rates by county to
decline, making it more difficult for any signal to achieve a high correlation
by county.

@@

As a further example, Figure~\ref{fig:bexar_compare} displays data from Bexar
County, Texas (which contains San Antonio) during July 2020. On July 16, 2020,
San Antonio reported 4,810 backlogged cases after reporting problems prevented
them from being reported over the past two weeks \cite{Palacios:2021}, resulting
in a clearly visible spike in the left-hand panel of the figure (case data from
JHU CSSE, smoothed using a 7-day trailing average). Meanwhile, Delphi's COVID
Trends and Impact Survey averaged around 350 responses per day in Bexar County
over the same time period, and was able to estimate the fraction of the
population who know someone in their local community with COVID-Like Illness
(CLI). As we can see in the right-hand panel of the figure, this signal was not
affected by Bexar County's reporting problems and, as shown in the last
subsection, it is (in general) highly correlated with case rates, providing an
alternate stream of data about COVID activity unaffected by backlogs. Similar
reporting problems have occurred in many jurisdictions across the United States,
and audits have regularly discovered misclassified or unreported cases and
deaths, making it valuable to cross-check against external sources not part of
the same reporting systems.

@@

Claims-based signals typically undergo heavy backfill as additional claims are
processed and errors are corrected; the median relative error between initial
reports and final values is over 20\% for such data, and only after roughly 35
days do estimates typically match finalized values within 5\%. However, the
systematic nature of this backfill, as illustrated in Figure~\ref{fig:dv_as_of},
suggests that statistical models could be fit (potentially separately for each
location) to estimate the final values from preliminary reports. On the other
hand, official public health reporting of COVID cases and deaths can be subject
to revision as death certificates are audited and backlogs cleared, resulting in
thousands of cases and deaths being added or removed.
Figure~\ref{fig:backfill-compare} shows relative differences between early
signal values, reported 10--90 days after the underlying events, and later
versions at least four months later. As the distribution of these revision
amounts is highly skewed, the figure plots the 95th percentile of relative
change, showing that reported deaths can have large revisions comparable to
revisions in claims-based signals. This backfill is much more difficult to
predict, and thus claims data and other sources may be a useful stand-in while
public health reports are aggregated and corrected.

\begin{figure}[t]
  \includegraphics[width=\columnwidth]{fig/backfill-compare.pdf}
  \caption{The 95th percentiles of relative changes between early reported
    values of key signals and final values reported much later. For each date
    between October 15, 2020 and April 15, 2021, the values for each state
    reported between 10 and 90 days later are compared to ``final'' versions
    recorded as of August 13, 2021, 120 days after April 15. Even officially
    reported case and death data can have large revisions 30--60 days or more
    after initial reporting; much of this is driven by individual large
    revisions affecting specific states and dates, rather than systematic
    changes affecting all states and dates.}
  \label{fig:backfill-compare}
\end{figure}

>>>>

Since July 2020, we have been regularly submitting short-term forecasts of
COVID-19 case and death incidence, at both the state and county levels, to the 
COVID-19 Forecast Hub \cite{ForecastHub}, with ``CMU-TimeSeries'' as the
team-model name.  The process of building, training, and deploying our
forecasting models leverages much of the infrastructure described in this paper
(such as the COVIDcast API's \textit{as of} feature), and some of our
forecasting systems rely on auxiliary indicators (such as survey-based and
claims-based COVID-like illness signals, which are described below).

@@

Besides tracking contemporaneous COVID activity, these and other indicators can
be used to improve forecasts of future COVID case trends, as investigated in our
companion paper \cite{McDonald:2021}.

@@

Figure~\ref{fig:geo_wise_correlation} shows the geo-wise correlations achieved
by the five signals and COVID case rates (from JHU CSSE, smoothed using a 7-day
trailing average), from April 15, 2020 to April 15, 2021. This calculation is
performed over all counties with at least 500 cumulative cases by the end of
this period, and at which all indicators are available (956 counties in total).
The large positive correlations suggest that these signals could be useful in
hotspot detection (identifying counties that have relatively high COVID
activity, at a given time). Somewhat surprisingly, the survey-based
CLI-in-community signal shows the strongest correlations for much of the time
period. This clearly demonstrates the value of a large-scale survey such as CTIS
for tracking symptoms and case trends, especially when other data is
unavailable.

Also notable is the fact that the correlations fluctuate over time in complex
ways; while some of this variation is likely due to changes in public behavior,
reporting and testing practices, and so on, some is also related to changes in
overall COVID-19 case trends. For example, comparing against
Figure~\ref{fig:time_trends_national}, we see that correlations drop in February
2021 for many of the signals, roughly matching the point in time when COVID
cases also sharply decline. This decline likely caused the heterogeneity in case
rates by county to decline, making it more difficult for any signal to achieve a
high geo-wise correlation.

@@

As a further example, Figure~\ref{fig:bexar_compare} displays data from Bexar
County, Texas (which contains San Antonio) during July 2020. On July 16, 2020,
San Antonio reported 4,810 backlogged cases after reporting problems prevented
them from being reported over the past two weeks \cite{Palacios:2021}, resulting
in a clearly visible spike in the left-hand panel of the figure (case data from
JHU CSSE, smoothed using a 7-day trailing average). Meanwhile, Delphi's COVID
Trends and Impact Survey averaged around 350 responses per day in Bexar County
over the same time period, and was able to estimate the fraction of the
population who know someone in their local community with COVID-Like Illness
(CLI). As we can see in the right-hand panel of the figure, this indicator was 
not affected by Bexar County's reporting problems and, as shown in the last 
subsection, it is (in general) highly correlated with case rates, providing an
alternate stream of data about COVID activity unaffected by backlogs. In
general, reporting problems have occurred in many jurisdictions across the
United States, and audits have regularly discovered misclassified or unreported
cases and deaths, making it valuable to cross-check against external sources not
part of the same reporting systems.

@@

Claims-based signals typically undergo heavy backfill as additional claims are
processed and errors are corrected; the median relative error between initial
reports and final values is over 20\% for such data, and only after roughly 35
days do estimates typically match finalized values within 5\%. However, the
systematic nature of this backfill, as illustrated in Figure~\ref{fig:dv_as_of},
suggests that statistical models could be fit (potentially separately for each
location) to estimate the final values from preliminary reports.

\begin{figure}[t]
  \includegraphics[width=\columnwidth]{fig/backfill-compare.pdf}
  \caption{The 95th percentiles of relative changes between early reported
    values of key signals and final values reported much later. For each date
    between October 15, 2020 and April 15, 2021, the values for each state
    reported between 10 and 90 days later are compared to ``final'' versions
    recorded as of August 13, 2021. Even officially reported case and death data
    can have large revisions 30--60 days or more after initial reporting; much
    of this is driven by individual large revisions affecting specific states
    and dates, rather than systematic changes affecting all states and dates.} 
  \label{fig:backfill-compare}
\end{figure}

Figure~\ref{fig:backfill-compare} shows relative differences between early
indicator values, reported 10--90 days after the underlying events, and later 
versions at least four months later. As the distribution of these revision
amounts is highly skewed, the figure plots the 95th percentile of relative
change, showing that reported deaths can incur large relative error in initial
values, comparable to that in claims-based signals.  However, for deaths, as
well as cases, the revision pattern is not very systematic, with large
corrections typically occuring at a sparse subset of locations and times
(e.g., due to audits or backlogs being cleared, which can result in thousands of
cases or deaths can be  added or removed all at once).  This backfill is much
more difficult to predict than that of claims data, thus the latter (and other 
sources) may be useful for nowcasting cases and deaths while public health 
reports are being aggregated and corrected.
